---
title: "Homework 2" 
author: "Hyun Suk Ryoo (hr2ee)"
fontsize: 12pt
geometry: margin=1in
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, highlight=FALSE)
```

## Problem 1
### A
```{r}
course <- list(Name="Gretchen Martinet", 
               Department= "Statistics", 
               Courses=c(2559,3080), 
               ActiveTeach2559= FALSE, 
               ActiveTeach3080=TRUE, 
               Enr2559=10, 
               Enr3080=c(90,90,88), 
               Days2559=c("Tuesday", "Thursday"), 
               Days3080=matrix(c("Monday", "Wednesday", "Tuesday", 
                                 "Thursday", "Tuesday", "Thursday"), 
                               nrow=3, ncol=2, byrow=TRUE))
print(course)
```

Made a list containing different dataframes

### B
```{r}
cap_3080 <- course$Enr3080
filled_3080 <- which(cap_3080 == 90)
print(filled_3080)
```
Sections 1 and 2 are filled at 90 students. This was done by going through the enrollment vector of 3080 and equating it to 90.

### C
```{r}
sum_reports <- sum(sum(course$Enr2559[course$ActiveTeach2559 == TRUE]) 
                   + 
                     sum(course$Enr3080[course$ActiveTeach3080 == TRUE]))
print(sum_reports)
```
I added all the numbers in the vector Enr3080 in the list of courses. 

## Problem 2
### A
```{r}
pnames <- c("Mercury", "Venus", "Earth", "Mars", 
            "Jupiter", "Saturn", "Uranus", "Neptune")
pdistance <-c(0.39, 0.72, 1, 1.52, 
              5.2, 9.54, 19.18, 30.06)
ptype <- c("terrestrial", "terrestrial", "terrestrial", "terrestrial", 
           "gas", "gas", "gas", "gas")
pdiameter <-c(0.382, 0.949, 1, 0.532, 
              11.209, 9.449, 4.007, 3.883)
protation <- c(58.64, -243.02, 1, 1.03, 
               0.41, 0.43, -0.72,0.67)
prings <-c(FALSE, FALSE, FALSE, FALSE, 
           TRUE, TRUE, TRUE, TRUE)
pmoons <-c("none", "none", "one", "more than one", 
           "more than one", "more than one", "more than one", "more than one")
planet_data <- data.frame(Name=pnames, distance=pdistance, 
                          type=ptype, diameter=pdiameter, 
                          rotation=protation, rings=prings, 
                          moons=pmoons)
print(planet_data)
```
I made several vectors and made it into a dataframe.

### B
```{r}
atleast2Au <- planet_data[which(planet_data$distance > 2), ]
print(atleast2Au)
```
I took which instances of the subset distance has a value greater than 2. Then I found the corresponding items in the original dataset that satisfies this condition. I then put it in a datafram called atlesat2Au

### C
```{r}
opposite_rotation <- which(planet_data$rotation <0)
print(opposite_rotation)
```
The two planets that are rotating the opposite way are Venus (2) and Uranus (7) with a distance of 0.72, 19.18 respectivly. 

### D
```{r}
diameter_bigger <- planet_data[which(planet_data$diameter > 1), ]
print(diameter_bigger[,c(1,3)])
```
I made data frame where the contents conditions were that they have a diameter greater than earth's, which is 1. I then proceeded to take the first and third columns to represent only the name and type.

### E
```{r}
morethanone_moon <- planet_data[which(planet_data$moons == "more than one"), ]
print(morethanone_moon[,c(2,3)])
```
I made a data frame where the contents conditions were that they have more than one moon. I then proceeded to take the second and third columns to isolate the distance and type of the original data frame. 

## Problem 3
### A
```{r}
USArrests2 <- USArrests[-3]
print(USArrests2[1:10,])
```

### B
```{r}
arrests_tot <- c(sum(USArrests2$Murder), 
                 sum(USArrests2$Assault), 
                 sum(USArrests2$Rape))
print(arrests_tot)
```
I took the sum of every crime and concated into a vector called arrests_tot

### C
```{r}
PropRape <- USArrests2$Rape / USArrests2$Assault
USArrests2$PropRape <- PropRape
print(USArrests2[1:10,])
```
I first took the proportions then I put them in as a vector. I then proceeded to make a dataframe.

### D
```{r}
above20 <- USArrests2[which(PropRape > 0.2), ]
states_20 <- row.names(above20[0])
print(states_20)
```
You can see from the dataframe above that the states that have a proportion higher than 20% ARE Hawaii, Iowa, Minnesota, Vermont, and Wisconsin. I used the row.names to get the names of the rows (states). I first took which proportions have higher than 20, then I took those indicies and set them as a parameter to find the names and details in the USArrests2 dataframe.

### E
```{r}
sorted_murder <- USArrests2[order(USArrests2$Murder, decreasing = TRUE),]
print(sorted_murder)
```
I used the order function with the parameter decreasing=TRUE  to sort the dataframe by descending murder counts. 

### F
```{r}
number_states <- sum(USArrests2$Murder > USArrests2$Rape)
print(number_states)
```
I compared the muder and rape cases to which there were no states.

## References
1. <https://stackoverflow.com/questions/7793295/
how-to-order-a-data-frame-by-one-descending-and-one-ascending-column>
2. <https://developmentality.wordpress.com/2010/02/12/
r-sorting-a-data-frame-by-the-contents-of-a-column/>
3. <https://stackoverflow.com/questions/1813550/count-of-entries-in-data-frame-in-r>
